Software Citation Principles

### 2016-03-07 sc-principles/software-citation-principles.tex
responding to comments in intro and principles

@@

 \acmnote{would like to add info on the problem we're trying to solve in academia to motivation}

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. \scnote{At least four styles (AMS, APA, Chicago, MLA) for
software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and widely used by
a community.
\item \textbf{Persistence}:
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available. Should we say anything about environments?}
\item \textbf{Specificity}:
Software citations should facilitate identification of and access to the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}

>>>>

There are many reasons this lack of both software citations in general and standard practices
for software citation are of concern:
\begin{itemize}
\item Software is a product of research, and by not citing it, we leave holes in the record of of scientific progress.
\item Academic researchers at all levels, including students, postdocs, faculty, and staff, should be
credited for the software products they develop and contribute to, particularly when those products enable or further research done by others.
\end{itemize}

\acmnote{would like to add info on the problem we're trying to solve in academia to motivation} 
\katznote{added list above - others can add more items if appropriate.}

@@

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Software should be cited whenever and wherever a research product (such as a paper) relies upon it, specifically, as part of the standard reference list for that research product.
\item \textbf{Credit and Attribution}:
Software citations should facilitate giving scholarly credit and normative
and legal attribution to all contributors to the software, recognizing
that a single style or mechanism of attribution may not be applicable to
all software. \scnote{At least four styles (AMS, APA, Chicago, MLA) for
software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item \textbf{Unique Identification}:
A software citation should include a method for identification that is
machine actionable, globally unique, interoperable, and widely used by
a community.
\item \textbf{Persistence}:
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Accessibility}:
Software citations should facilitate access to the software itself and to such associated metadata, documentation, data, and other materials as are necessary for both humans and machines to make informed use of the referenced software.\dmnote{Should provide a location where the software is available. Should we say anything about environments?} \katznote{If we provide access, I think that includes the location.  I don't think we need to say anything about environments here, since it is in the next principles (platforms).  Daniel, if this is ok with you, please comment out this set of comments}
\item \textbf{Specificity}:
Software citations should facilitate identification of and access to the specific version of software that was used.  Software identification should be as specific as necessary, such as using version numbers, revision numbers, or variants such as platforms.
%\item \textbf{Interoperability and Flexibility}:
%Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
\dmnote{ Anything missing? Configurations? Interactive software? Screenshots? \ldots ?}

\katznote{perhaps we should have some discussion of these here, so that we could include Sou-Cheng's note and others?  I would like to have the principles be as brief as possible, but in some cases, more may be needed.  Or perhaps that comment goes into the next section?}
